---
title: "Untitled"
output: html_document
---
# INSTALLING PACKAGES 
```{r}
install.packages("tidyverse")
library(tidyverse)
```

```{r}
install.packages("VIM")
library(VIM)
```

```{r}
install.packages("dplyr")
library(dplyr)
```

# Renaming datasets
```{r}
energy = energy_dataset
weather = weather_features
```

# 1. Summary of sepearte datasets
```{r}
# Checking Datasets
# Summaries for energy dataset
summary(energy) 
str(energy)
# Summaries for weather_features dataset
summary(weather) 
str(weather)
```


# 2.Combining datasets using DPLYR #
```{r}
energy_weather <- inner_join(weather_features,energy,by=c("dt_iso" ="time"),suffix=c("_W","_E"))
```

# 3. Loading Data #
```{r}
str(energy_weather)
```


# 4. Renaming Columns
```{r}
names(energy_weather)[names(energy_weather) == "dt_iso"] <- "Time"
```

# 5. Checking for missing values in combined dataset
```{r}
#identify missing values
summary(energy_weather)

#Finding number of NAs
energy_weather.NACount <- apply(is.na(energy_weather), 2, sum )
energy_weather.NA.PER <- energy_weather.NACount / dim(energy_weather)[1] * 100
energy_weather.NA.PER
```

# 6. Removing columns 
# The following columns are being removed because of the data only consisting of Zero values. 
# Additionally, the columns removed were the columns that included missing values 
```{r}
energy_weather = select(energy_weather, -c(`generation fossil coal-derived gas`, `generation fossil oil shale`, `generation fossil peat`, `generation geothermal`, `generation hydro pumped storage aggregated`, `generation marine`, `generation wind offshore`, `forecast wind offshore eday ahead`))
```


#7. Checking Missingness
```{r}
#Finding Missing Values by creating useful functions
# SumNA function
sumNa <- function(x){
	ty= is.na(x)
	yt= sum(ty)
	return(yt)
}

# Checking for missingness
apply(energy_weather, 2, sumNa)

# Omitting NA values
energy_weather = na.omit(energy_weather)
```



# 8. Handling duplicates
```{r}
# Finding the number of non-duplicate values
energy_weather_nodup <- unique(energy_weather)

# Removing duplicate values
energy_weather <- energy_weather[!duplicated(energy_weather),]
```

# Number of NA
```{r}
energy_weather.NACount <- apply(is.na(energy_weather), 2, sum )
energy_weather.NACount
```



# 9. Checking for outliers
```{r}
energy_weather_boxplot <- boxplot(energy_weather$) 
min_ _outlier <- min(energy_weather_boxplot$out)
clean_energy_weather <- energy_weather[energy_weather$ ...  < min_  _outlier, ]
```




















