---
title: "GROUPPP"
output: html_document
---
1. Load data 
```{r}
str(netflix_version6)
```

2. Removing Columns 

# We removed show Seasons as the seasons are also included in duration
# We removed Genre because it was similar to listen in, as well as having too many NA's
# We removed Show Id as it was the id for imbd which has nothing to do with netflix
```{r}
# Removing columns that are not necassary for the data 
netflix_version6 = select(netflix_version6, -c(show_id_imdb, Seasons, Genre,))
view(netflix_version6)
```



3. Checking for missing values
# We found missing values in the data and removed NA's in Director_original to "No Director"
# We do not think we should remove NA's in country as we believe it would be good for EDA to play around with.
```{r}
#identify missing values
summary(netflix_version6)

#Finding number of NAs
netflix.NACount <- apply(is.na(netflix_version6), 2, sum )
netflix.NA.PER <- netflix.NACount / dim(netflix_version6)[1] * 100
netflix.NA.PER

#Removing NAs by column
netflix_version6$director_orginal[is.na(netflix_version6$director_orginal)] <- "No Director"
view(netflix_version6)

```


4. Impute missing values to median value for IMDB rating
```{r}
# Finding median rating value
median_rating = median(netflix_version6$rating_imdb)
median_rating

# Changing all Zero values to the median value
netflix_version6$rating_imdb[netflix_version6$rating_imdb == "0"] <- "6.8"
```


5. Handling duplicates 
```{r}
# Finding the number of non-duplcate values
length(unique(netflix_version6$description_orginal))

# Removing duplicate values
Netflix_New <- netflix_version6[!duplicated(netflix_version6$description_orginal),]
```


6. Fixing Netflix Original to binary
#  CHARACTER FOR NETFLIX ORIGINAL CANNOT BE CHANGED!!!

```{r}
Netflix_New$Original.Network[Netflix_New$Original.Network == "Netflix"] <- 1
Netflix_New$Original.Network[is.na(Netflix_New$Original.Network)] <- 0
Netflix_New$Original.Network[Netflix_New$Original.Network == "abc"] <- 0
Netflix_New$Original.Network[Netflix_New$Original.Network == "Fox"] <- 0
Netflix_New$Original.Network[Netflix_New$Original.Network == "showcase"] <- 0
Netflix_New$Original.Network[Netflix_New$Original.Network == "Channel 4"] <- 0
```


7. Handling Outliers

# We believe we do not need to identify any outliers as the rating can only range from 0-10. If the rating exceeded 10 or was below 0, then that would be considered an outlier. 
```{r}

Netflix_New$rating_imdb <- as.numeric(Netflix_New$rating_imdb)
```

```{r}
# Remove Instances with Outlier Values 
netflix_New_boxplot <- boxplot(Netflix_New$rating_imdb) 
min_imdb_outlier <- min(netflix_New_boxplot$out)
clean_netflix <- Netflix_New[Netflix_New$rating_imdb < min_imdb_outlier, ]

```


8. Fixing Date Format?????????
# We tried to change the format of the date in excel as we couldnt change it here. That was also not possible. 
```{r}
# Convert date information in format 'dd/mm/yy'
class(netflix_version6$date_added_orginal)

#Convert character string column to date

```


9. Renaming Data 
```{r}
names(Netflix_New)[names(Netflix_New) == "title"] <- "Title"
names(Netflix_New)[names(Netflix_New) == "type_orginal"] <- "Type"
names(Netflix_New)[names(Netflix_New) == "director_orginal"] <- "Director"
names(Netflix_New)[names(Netflix_New) == "country_orginal"] <- "Country"
names(Netflix_New)[names(Netflix_New) == "date_added_orginal"] <- "Date_Added"
names(Netflix_New)[names(Netflix_New) == "release_year_orginal"] <- "Release_Year"
names(Netflix_New)[names(Netflix_New) == "listed_in_orginal"] <- "Genre"
names(Netflix_New)[names(Netflix_New) == "rating_imdb"] <- "IMDB_Rating"
names(Netflix_New)[names(Netflix_New) == "description_orginal"] <- "Description"
names(Netflix_New)[names(Netflix_New) == "Original.Network"] <- "Netflix_Original"
names(Netflix_New)[names(Netflix_New) == "duration"] <- "Duration"
view(Netflix_New)

```




























